BOLD Parameter Estimation Using Particle Filters

### 2010-08-14 doc/thesis/results.tex
more on the results. Updated the HRF and applied pierces recommendations

\chapter{Simulation}
\label{sec:SimulationResults}
Two levels of simulation are discussed in this section. First the simulation of single time series
and second, a slice of simulated FMRI time series. The single time series tests
are necessary to investigate the power of particle filters in 
identifying the BOLD model in a noisy environment. If the particle filter is unable
to get back to the \emph{true} time-series in spite of noise then the particle filter
will not be useful. Additionally, determination of reasonable parameters from the
time-series is also desirable. An entire slice is then simulated using 
Physics-Oriented Simulated Scanner for Understanding MRI (POSSUM) which
models noise more realistically. POSSUM also provides insight into the 
applicability of particle filters in modeling the BOLD 
signal on a large scale. Both sorts of tests are usefull in determining the
consistency of the particle filter. For the single voxel tests, each test will
actually be performed eleven different times, each with a new noise realization.
If runs fail to have consistent estimates of the underlying BOLD signal, then
this would indicate a large variance error in the model. Similarly, in the
Possum simulation there are actually only four discrete parameter sets driving the time-series.
Nevertheless each voxel will have a different tissue composition and a different
noise realization. Therefore the Possum simulation should also provide a good
test of the consistency of the BOLD model.

\section{Single Time Series}
\label{sec:Single Voxel Simulation}
Given the state-space equations for the BOLD signal, simulating a single time 
series was relatively straight forward. After generating a true signal,
identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps were added to this true signal. Finally a 
carrier level was added, since BOLD is typically
measured as a \% difference from the baseline. The particle filter
algorithm will immediately remove this by calculating the \% difference, 
but adding a carrier level meant that the exact same algorithm used 
for simulated data could be used for the real data.

Once this noisy simulated time series was generated, the particle filter algorithm
 was run on the single time series image. Here I include four sets of tests 
to determine the power of the particle filter in modeling. The first two were
used to demonstrate the ability of the particle filter to find the most likely
set of parameters/state variables over the course of the run. The second two 
tests investigated the problem of false-positives. As the first
round of tests show, given that there is an underlying BOLD signal, 
the particle filter is excellent at finding the most probable cause of 
the observed signal. However, as discussed in the false positive tests,
even if when there was no underlying signal, the particle filter still determined
the most likely BOLD signal; which was not 0. Thus, it was
important to investigate methods of determining false positives using 
other methods.

%LOW NOISE SECTION, with a signal
\subsection{Signal with Low Noise}
\label{sec:SimLowNoise}
%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
%parameters .
%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
%thus the particle filter was not centered over the correct values. 
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal.}
\label{fig:LowNoiseRealization}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_lownoise}
\caption{A comparison of the preprocessed signals for the low noise case. This is the
noisy input to the actual particle filter algorithm.}
\label{fig:PreprocessedLowNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_lownoise}
\caption{A comparison of the fitted signals for the low noise case.}
\label{fig:FitComparisonLowNoise}
\end{figure}

To begin the single voxel simulation; I generated a signal using the following parameters:
$\{\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
These same parameters were also used in the high noise case. The noise realizations were
then generated based on a measurement noise ($\sigma_y$) of $.001$ and a drift standard deviation
($\sigma_x$) of $.0005$. The measurement noise as well as the steps of the drift
were taken to be Gaussian.
For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
The actual signal delivered into the particle filter
was the result of preprocessing to remove drift, as described in 
\autoref{sec:Methods Preprocessing}. The bias introduced into the signal by preprocessing 
may have some effect on the resulting fit; thus the preprocessed signal is compared
to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
The data shows that the spline struggles to match the signal near the end, although 
overall, the preprocessing seems to have been successful at removing trends. While finding
a better preprocessing algorithm could be useful, non will ever be perfect. In this case
the cause of the problem was the final stimuli that pushed the final median well above the base
level. The fitting lines to the true BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
 In several locations the output
of the particle filter looks like a filtered version of the input. For instance toward the
end, the estimates stay flat in spite of the preprocessed data drifting off. By
this point, the algorithm has converged sufficiently to prevent inexplicable movement.
A similar circumstance occurs at 100 seconds. Note that bias introduced by pre-processing 
results in a slightly larger peak signal. The final parameter sets are shown in 
\autoref{tab:LowNoiseResults}. Note that throughout the results, unless otherwise specified,
the term \emph{residual} will be defined as the square root of the mean squared residual, 
\begin{equation}
\text{Residual} = \sqrt{\sum (\hat{y}_k - y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the output
sample at time $k$. Note the subtle distinction between this and what will be 
termed the \emph{error} which is defines as 
\begin{equation}
\text{Error} = \sqrt{\sum (\hat{y}_k - Y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
signal. Throughout this section and the next, the terms error and residual
will be referencing these values (for a particular timeseries).

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSR}$ &$\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.009876  \\
1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.009932  \\
1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.009680  \\
1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.009120  \\
1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.009713  \\
1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.009485  \\
1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.010219  \\
1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.010080  \\
1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.010329  \\
1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.010015  \\
1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.009505  \\
\hline                                                                           
1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.009814 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
last is mean over the 11 runs. The $\sqrt{MSR}$ is residual and $\sqrt{MSE}$ is the error.}
\label{tab:LowNoiseResults} 
\end{table}

@@

\subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $V$ of the first run, low noise simulation.]
{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise2}}\\
\end{figure}

\begin{figure}
\subfigure[Converging histogram for $Q$, $S$, $F$, and $BOLD$ of the first run, low noise simulation.]
{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise3}}
\label{fig:LowNoiseHist}
\end{figure}

There are a few important results in the final parameter estimates (which 
are the mean of the particle filter's posterior distribution). First, the time constants vary 
greatly across
runs with different noise realizations, yet the sum of the individual time constants
($\tau_f$, $\tau_s$ and $\tau_0$) seems to be more consistent. In general the 
time constants fall short of the true time constant. This could be a limitation
based on the prior distribution (which notably has an initial mean below the true values) 
or it could be other parameters are compensating. Its also possible that output is insensitive
to differences in the time constants; although varying the activation durations could induce
better estimates.  
The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
demonstrates the migration of parameters to match the estimates. In spite of the
rather wide final parameter variances, the time series'
generated from the estimated mean seems to match the correct signal very well 
(\autoref{fig:FitComparisonLowNoise}).

@@

%HIGH NOISE SECTION, with a signal
\subsection{High Noise}
\label{sec:HighNoise}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/realization_highnoise}
\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
\label{fig:HighNoiseRealization}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/preprocessed_highnoise}
\caption{A comparison of the preprocessed signals for the high noise case.}
\label{fig:PreprocessedHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise}
\caption{A comparison of the fitted signals for the high noise case.}
\label{fig:FitComparisonHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/highnoise_56_noise}
\caption{Two particular preprocessed noise realizations for the high noise case.}
\label{fig:NoiseComparisonJustTwo}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise_just2}
\caption{The results for the noise realizations shown in \autoref{fig:NoiseComparisonJustTwo}.}
\label{fig:FitComparisonHighNoiseJust2}
\end{figure}

For the high noise simulation, the exact same procedure was performed as with the lower
noise simulation except that $\sigma_y$ and $\sigma_x$ were set to $.01$ and $.005$,
respectively. This is an order of magnitude higher than the previous tests, and indeed the
noise appears to dominate the output, as seen in \autoref{fig:HighNoiseRealization}.  
As before, the results of preprocessing are helpful in identifying the reason for
bias in the estimated output signal; thus the preprocessed signals are shown in 
\autoref{fig:PreprocessedHighNoise}. The results of the particle filter
for each of the eleven runs are shown in \autoref{fig:FitComparisonHighNoise}. 
As before, the preprocessing led the algorithm
to somewhat higher activation levels, and it would appear that the subtleties of
different time constants are being lost due to the noise. This is easily seen
by observing the post stimulus undershoots; many of the estimated time series
have no post-stimulus undershoot but rather decay directly to the base level.

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSR}$  & $\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.1900 & 0.2349 & 0.4223 & 0.128  & 1.0147 & 2.4779 & 1.1168 & 4.6826 &0.01406 &0.01573   \\
0.9721 & 0.2190 & 0.3051 & 0.061  & 0.5780 & 1.9960 & 3.4613 & 3.5461 &0.01373 &0.01378   \\
1.5795 & 0.1415 & 0.3380 & 0.1089 & 0.5843 & 2.1247 & 1.7834 & 4.2885 &0.01275 &0.01577   \\
1.1094 & 0.2374 & 0.5349 & 0.0351 & 1.2186 & 3.0736 & 2.3504 & 5.4016 &0.01673 &0.01154    \\
1.1071 & 0.2753 & 0.3365 & 0.0316 & 1.5057 & 2.6518 & 4.1910 & 5.2646 &0.01370 &0.01222   \\
0.5803 & 0.4793 & 0.4135 & 0.1189 & 0.9756 & 3.6902 & 1.0008 & 5.2461 &0.01150 &0.01316   \\
\rowcolor[rgb]{.9,.5,.5}
1.2952 & 0.2596 & 0.2756 & 0.2595 & 1.7026 & 2.8458 & 0.6617 & 5.8436 &0.01555 &0.01790   \\
\rowcolor[rgb]{.5,.5,.9}
1.5185 & 0.2199 & 0.2835 & 0.0742 & 0.8882 & 3.0771 & 1.7393 & 5.4838 &0.01205 &0.01246   \\
0.6874 & 0.3283 & 0.3979 & 0.1561 & 1.0778 & 3.1158 & 0.6643 & 4.8810 &0.01510 &0.01258   \\
1.0170 & 0.285  & 0.3474 & 0.0567 & 1.5877 & 2.6516 & 2.2852 & 5.2563 &0.01249 &0.01343   \\
0.9925 & 0.298  & 0.3221 & 0.2094 & 0.4276 & 2.2108 & 1.0167 & 3.6308 &0.01217 &0.01506   \\
\hline                                                                          
1.0954 & 0.2708 & 0.3615 & 0.1126 & 1.0510 & 2.7196 & 1.8428 & 4.8659 &0.01362 &0.01397\\
\hline 
\end{tabular}
\caption{Estimated Parameters on 10 different runs with high noise. First row is the true parameters.
The red row is Run 1 and the blue row is Run 2, as named in  \autoref{fig:NoiseComparisonJustTwo}
and \autoref{fig:FitComparisonHighNoiseJust2}}
\label{tab:HighNoiseResults} 
\end{table}

It is interesting to consider how the preprocessing and noise may effect the
parameters of the fitting model. Consider two runs, shown in 
\autoref{fig:NoiseComparisonJustTwo}, out of the previous eleven. 
Run 1 has higher peaks, in spite of the fact that both signals are based
on the same underlying signal. There also appears
to be more drift than the 20 measurements per knot could fit, which explains run 1's
prolonged increase at 170 seconds in. Interestingly, run 1 and run 2 have different
aspects of the signal that match better. Run 2 had a much better match to the 
peaks, when compared to the true signal, yet run 1 matched the post-stimulus
undershoot better. Its worth noting however, that the post-stimulus undershoots
here are much shorted than the observed prolonged post-stimulus undershoot
discussed in \autoref{sec:Post Stimulus Undershoot}. 
Despite the rather significant amount of noise, the ultimate results are 
actually rather good. \autoref{tab:HighNoiseResults} shows
the square-root mean-squared-error for all eleven runs, and highlights the two runs analyzed 
in \autoref{fig:ConvergenceRuns1} and \autoref{fig:ConvergenceRuns2}.

\begin{figure}
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 1]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_1}}\\
\end{figure}
\begin{figure}
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 1] 
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_2}}\\
\end{figure}
\begin{figure}
\subfigure[$Q$, $S$, $F$, $BOLD$, Run 1 ]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_3}}
\caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
\label{fig:ConvergenceRuns1}
\end{figure}

>>>>

\chapter{Simulation}
\label{sec:SimulationResults}
Two levels of simulation are discussed in this section; first of single voxels,
and second of a single slice (64x64 voxels).  The single time series tests
investigate the ability of particle filters to estimate the BOLD model 
in a noisy environment. If the particle filter were unable
to identify the \emph{true} time-series in spite of noise, then it 
would not be a good choice for identifying parameters. Single voxel tests were
performed eleven different times, each with a new noise realization.
If runs failed to have consistent estimates of the underlying BOLD signal, then
that would indicate a significant variance error in the model.

The slice simulation was performed using 
Physics-Oriented Simulated Scanner for Understanding MRI (POSSUM) which
models noise realistically. POSSUM demonstrates the 
particle filter modeling the BOLD signal on a large scale. 
In the POSSUM simulation there were four
discrete parameter sets driving the timeseries', however each voxel had a 
different tissue composition and a different
noise realization. Therefore the POSSUM simulation was 
a good test of the applicability of particle filters for performing
whole brain analysis.

\section{Single Time Series}
\label{sec:Single Voxel Simulation}
Given the state-space equations for the BOLD signal, simulating a single time 
series was straight forward. After generating a true signal,
identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps were added to this true signal. Finally a 
carrier level was added, since BOLD is typically
measured as a \% difference from the baseline. The particle filter
algorithm immediately removes this by calculating the \% difference, 
but adding a carrier level meant that the exact same algorithm used 
for simulated data could be used for the real data.

Once this noisy simulated time series was generated, the particle filter algorithm
 was run on the single time series image. Five sets of tests were used 
to determine the power of the particle filter in modeling. The first two (\autoref{sec:SimLowNoise}
and \autoref{sec:SimHighNoise}) were
used to demonstrate the ability of the particle filter to find the most likely
set of parameters/state variables over the course of the run. The second two 
tests (\autoref{sec:PureNoiseLowMag} and \autoref{sec:PureNoiseHighMag}) 
investigated the problem of false-positives. As the first
round of tests will show, given that there is an underlying BOLD signal, 
the particle filter is excellent at finding the most probable cause of 
the observed signal. As discussed in \autoref{sec:PureNoiseLowMag} and 
\autoref{sec:PureNoiseHighMag}, even when there was no underlying signal, 
the particle filter still converged. Thus, it was 
important to investigate methods of identifying false positives.

The final single-voxel test was used to determine the effectiveness of the
particle filter for a longer FMRI session. The final covariance and correlation
matrices are also discussed.

%LOW NOISE SECTION, with a signal
\subsection{Signal with Low Noise}
\label{sec:SimLowNoise}
%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
%parameters .
%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
%thus the particle filter was not centered over the correct values. 
\begin{figure}
\centering
\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal.}
\label{fig:LowNoiseRealization}
\end{figure}

\begin{figure}
\centering
\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/preprocessed_lownoise}
\caption{A comparison of the preprocessed signals for the low noise case. The
noisy input to the actual particle filter algorithm.}
\label{fig:PreprocessedLowNoise}
\end{figure}

\begin{figure}
\centering
\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/comparison_lownoise}
\caption{A comparison of the fitted signals for the low noise case.}
\label{fig:FitComparisonLowNoise}
\end{figure}

To begin the single voxel simulation; I generated a signal using the following parameters:
$\{\tau_0 = 1.45, \alpha = 0.3, E_0 = 0.47, V_0 = 0.044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
These same parameters were used for all the tests in this chapter. The noise realizations were
then generated based on a measurement noise ($\sigma_y$) of $0.001$ and a drift standard deviation
($\sigma_x$) of $0.0005$. The measurement noise as well as the steps of the drift
were taken to be Gaussian.
For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
The actual signal delivered into the particle filter
was the result of preprocessing to remove drift, as described in 
\autoref{sec:Methods Preprocessing}. The bias introduced into the signal by preprocessing 
could have some effect on the resulting fit; thus the preprocessed signal is compared
to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
The data shows that the spline struggles to match the signal near the end, although 
overall, the preprocessing seems to have been successful at removing trends.  In this case
the cause of the problem was the final stimuli that pushed the final median well above the base
level. The estimated BOLD signal is compared to the true BOLD signal in \autoref{fig:FitComparisonLowNoise}.
At several time points the particle filter successfully filtered the input. 
For instance, in the last 30 seconds
the estimates stay flat in spite of the preprocessed data drifting off. By
this point, the algorithm has converged sufficiently to prevent inexplicable movement.
A similar circumstance occurs at 100 seconds in. Note that bias introduced by pre-processing 
results in a slightly larger peak signal. The final parameter sets are shown in 
\autoref{tab:LowNoiseResults}. Note that throughout the results, unless otherwise specified,
the term \emph{residual} will be defined as the square root of the mean squared residual, 
\begin{equation}
\text{Residual} = \sqrt{\sum (\hat{y}_k - y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the preprocessed output
sampled at time $k$. Note the subtle distinction between this and what will be 
termed the \emph{error} which is defines as 
\begin{equation}
\text{Error} = \sqrt{\sum (\hat{y}_k - Y_k)^2}
\end{equation}
where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
signal. Throughout this section and the next, the terms error and residual
will be referencing these values.

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Residual & Error\\
\hline 
\rowcolor[gray]{.8}
1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.009876  \\
1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.009932  \\
1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.009680  \\
1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.009120  \\
1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.009713  \\
1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.009485  \\
1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.010219  \\
1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.010080  \\
1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.010329  \\
1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.010015  \\
1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.009505  \\
\hline                                                                           
1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.009814 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise. First row is the true value, and last is the average.}
\label{tab:LowNoiseResults} 
\end{table}

@@

\subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $v$ of the first run, low noise simulation.]
{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise2}}\\
\end{figure}

\begin{figure}
\subfigure[Converging histogram for $q$, $s$, $f$, and $BOLD$ of the first run, low noise simulation.]
{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise3}}
\label{fig:LowNoiseHist}
\end{figure}

There are a few important results in the final parameter estimates (which 
are the mean of the particle filter's posterior distribution). First, the time constants vary 
greatly across
runs with different noise realizations, yet the sum of the individual time constants
($\tau_f$, $\tau_s$ and $\tau_0$) seems to be more consistent. In general the 
time constants fall short of the true time constant. This could be a limitation
based on the prior distribution (which notably has an initial mean below the true values) 
or it could be other parameters are compensating. It is also possible that output is insensitive
to differences in the time constants.  
The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
demonstrates the migration of parameters to match the estimates. In spite of the
wide final parameter variances, the time series'
generated from the estimated mean match the correct signal very well 
(\autoref{fig:FitComparisonLowNoise}).

@@

%HIGH NOISE SECTION, with a signal
\subsection{High Noise}
\label{sec:SimHighNoise}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/realization_highnoise}
\caption{Test Signals with high noise ($\sigma_x = 0.01, \sigma_y=0.005$) compared to the clean signal.}
\label{fig:HighNoiseRealization}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/preprocessed_highnoise}
\caption{A comparison of the preprocessed signals for the high noise case.}
\label{fig:PreprocessedHighNoise}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/comparison_highnoise}
\caption{A comparison of the fitted signals for the high noise case.}
\label{fig:FitComparisonHighNoise}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/highnoise_56_noise}
\caption{Two particular preprocessed noise realizations for the high noise case.}
\label{fig:NoiseComparisonJustTwo}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/comparison_highnoise_just2}
\caption{The results for the noise realizations shown in \autoref{fig:NoiseComparisonJustTwo}.}
\label{fig:FitComparisonHighNoiseJust2}
\end{figure}

For the high noise simulation, the exact same procedure was followed as in \autoref{sec:SimLowNoise}
except that $\sigma_y$ and $\sigma_x$ were set to $0.01$ and $0.005$,
respectively. This is an order of magnitude higher than the previous tests, and indeed the
noise appears to dominate the output, as seen in \autoref{fig:HighNoiseRealization}.  
As before, the results of preprocessing are helpful in identifying the reason for
bias in the estimated output signal; thus the preprocessed signals are shown in 
\autoref{fig:PreprocessedHighNoise}. The results of the particle filter
for each of the eleven runs are shown in \autoref{fig:FitComparisonHighNoise}. 
The preprocessing again led the estimate
to somewhat higher activation levels, and it would appear that the subtleties of
different time constants are being lost due to the noise.

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Residual  & Error\\
\hline 
\rowcolor[gray]{.8}
1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.1900 & 0.2349 & 0.4223 & 0.128  & 1.0147 & 2.4779 & 1.1168 & 4.6826 &0.01406 &0.01573   \\
0.9721 & 0.2190 & 0.3051 & 0.061  & 0.5780 & 1.9960 & 3.4613 & 3.5461 &0.01373 &0.01378   \\
1.5795 & 0.1415 & 0.3380 & 0.1089 & 0.5843 & 2.1247 & 1.7834 & 4.2885 &0.01275 &0.01577   \\
1.1094 & 0.2374 & 0.5349 & 0.0351 & 1.2186 & 3.0736 & 2.3504 & 5.4016 &0.01673 &0.01154    \\
1.1071 & 0.2753 & 0.3365 & 0.0316 & 1.5057 & 2.6518 & 4.1910 & 5.2646 &0.01370 &0.01222   \\
0.5803 & 0.4793 & 0.4135 & 0.1189 & 0.9756 & 3.6902 & 1.0008 & 5.2461 &0.01150 &0.01316   \\
\rowcolor[rgb]{.9,.5,.5}
1.2952 & 0.2596 & 0.2756 & 0.2595 & 1.7026 & 2.8458 & 0.6617 & 5.8436 &0.01555 &0.01790   \\
\rowcolor[rgb]{.5,.5,.9}
1.5185 & 0.2199 & 0.2835 & 0.0742 & 0.8882 & 3.0771 & 1.7393 & 5.4838 &0.01205 &0.01246   \\
0.6874 & 0.3283 & 0.3979 & 0.1561 & 1.0778 & 3.1158 & 0.6643 & 4.8810 &0.01510 &0.01258   \\
1.0170 & 0.285  & 0.3474 & 0.0567 & 1.5877 & 2.6516 & 2.2852 & 5.2563 &0.01249 &0.01343   \\
0.9925 & 0.298  & 0.3221 & 0.2094 & 0.4276 & 2.2108 & 1.0167 & 3.6308 &0.01217 &0.01506   \\
\hline                                                                          
1.0954 & 0.2708 & 0.3615 & 0.1126 & 1.0510 & 2.7196 & 1.8428 & 4.8659 &0.01362 &0.01397\\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with high noise. First row contains the true parameters,
last row contains the mean.
The red row is Run 1 and the blue row is Run 2 from  \autoref{fig:NoiseComparisonJustTwo}
and \autoref{fig:FitComparisonHighNoiseJust2}, respectively.}
\label{tab:HighNoiseResults} 
\end{table}

Two runs are shown in more detail in \autoref{fig:NoiseComparisonJustTwo}.
Run 1 has higher peaks, in spite of the fact that both signals are based
on the same underlying signal. There also appears
to be more drift than the 20 measurements per knot could fit, which explains run 1's
prolonged increase at 170 seconds. Interestingly, run 1 and run 2 emphasize
different aspects of the signal. Run 2 had a much better match to the 
peaks, when compared to the true signal, yet run 1 matched the post-stimulus
undershoot better. Its worth noting that the post-stimulus undershoots
here are much shorter than the observed prolonged post-stimulus undershoot
discussed in \autoref{sec:Post Stimulus Undershoot}. 
Despite the rather significant amount of noise, the ultimate results are 
actually quite good. \autoref{tab:HighNoiseResults} shows
the error  for all eleven runs, and highlights the two runs analyzed 
in \autoref{fig:ConvergenceRuns1} and \autoref{fig:ConvergenceRuns2}.

\begin{figure}
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 1]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_1}}\\
\end{figure}

\begin{figure}
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$, Run 1] 
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_2}}\\
\end{figure}

\begin{figure}
\subfigure[$q$, $s$, $f$, $BOLD$, Run 1 ]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_3}}
\caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
\label{fig:ConvergenceRuns1}
\end{figure}

\begin{figure}
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 2]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_1}}\\
\end{figure}
\begin{figure}
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$, Run 2] 
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_2}}\\
\end{figure}
\begin{figure}
\subfigure[$q$, $s$, $f$, $BOLD$, Run 2 ]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_3}}
\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
\label{fig:ConvergenceRuns2}
\end{figure}
